### 2020-01-16 ISSTA/implementation.tex
small shrink

Note that TSTL provides explicit support for exploring the real-world sources of
nondeterminism described in Section \ref{sec:sources}, by supporting
execution in a new process and allowing
control over timing changes in re-execution, by specifying a delay to
introduce between actions.  Changing the timing of
steps is very effective when exact timing is somehow involved in
results, and often effective in exposing concurrency-related
nondeterminism, since the delay will frequently change the
interleaving of threads.

@@

First-class vertical nondeterminism checks are currently
limited to failures:  {\tt --checkFailureDeterminism} causes the TSTL
compiler to emit a harness where every action that causes an \emph{expected}
exception to be raised is repeated to ensure the same exception is
still raised.  In terms of the formalism, if execution of action
$\alpha_i$ raises an exception, then there are two possibilities:

\begin{enumerate}
\item If the TSTL harness does not explicitly allow the raised
  exception, then the test fails.  There is no check for vertical
  nondeterminism in this case, since the behavior of the system is
  incorrect whether or not the behavior is idempotent.
  \item If the harness does allow exception $e$ (e.g., as with {\tt
      KeyError} and {\tt ValueError} in Figure \ref{fig:simple}), then
    $F_i$ is true, and we repeat $\alpha_i$ in order to ensure that
    $F_{i+1}$ is also true.
    \end{enumerate}

When a failure is indicated by some means other than raising an
=exception, TSTL users must express this by explicitly writing
a wrapper that raises an expected exception on failure.

>>>>

Note that TSTL provides explicit support for exploring the real-world sources of
nondeterminism described in Section \ref{sec:sources}, by supporting
execution in a new process and allowing
control over timing changes.  Changing the timing of
steps is also effective in exposing concurrency-related
nondeterminism, since the delay will frequently change the
interleaving of threads.

@@

First-class vertical nondeterminism checks are currently
limited to failures.  
%{\tt --checkFailureDeterminism} causes the TSTL
%compiler to emit a harness where every action that causes an \emph{expected}
%exception to be raised is repeated to ensure the same exception is
still raised.
%In terms of the formalism, if execution of action
$\alpha_i$ raises an exception that does not, itself, indicate a test failure, then 
    $F_i$ is true, and TSTL repeats $\alpha_i$ in order to ensure that
    $F_{i+1}$ is also true.
    %\end{enumerate}

%When a failure is indicated by some means other than raising an
%exception, TSTL users must express this by explicitly writing
%a wrapper that raises an expected exception on failure.
